% This file is part of the i10 thesis template developed and used by the
% Media Computing Group at RWTH Aachen University.
% The current version of this template can be obtained at
% <http://www.media.informatik.rwth-aachen.de/karrer.html>.

\chapter{Related work}
\label{relatedwork}
\section{Knowledge and Task Mining}
In the computer scientific field of knowledge and task mining is no new subject.\mnote{Context-Aware Systems} With the rise of mobile computing devices the term \ac{CAS} was created. The meaning and definition are disputed. First publications referred to a user's location: in different places usually different contextual parameters are relevant. For example a diver that is ascending from deep water has to be made aware of resting times before emerging to the surface.Another example was the \textit{Active Badge Location System} in 1992 that detected the whereabouts of a person and in order to forward relevant phone calls to telephones close targeted person (\cite{want1992active}). Such systems adapt not only to the location but also to other relevant and changing parameters in the surroundings (\cite{schilit1994context}). This definition was widened in 1998 where context was referred to not only the computer accessible parameters of the surroundings but also the emotional state, focus of attention, date and time as well as people in the environment (\cite{dey1998context}). The new aspect of internal parameters like focus of attention was then referred to a further elaboration of the definition: the internal (logical) and external context: Internal context parameters are specified by the user in interaction with the computer like goals, tasks, work context, business processes and emotional state. External parameters are usually measured by hardware sensors, i.e. location, light, sound, movement, temperature, pressure etc. (\cite{hofer2003context}). The contextual parameters can be grouped into four categories: identity (marked by a unique identifier), the location (an entity’s position), activity (status, meaning the intrinsic properties of an entity, e.g., temperature and lightning for a room, processes running currently on a device etc.) and time (timestamps, \cite{dey2001conceptual}). An example of the use of internal data for extracting context is the \textit{Watson Project} \cite{budzik2000user}. Here the focus is shifted for collecting contextual information from user interaction with the computer in order to proactively support the user. Proactivity is a term that  originates in organizational psychology and describes the ability of workers to not react to situations, but sense upcoming situational changes in advance and take control (\cite{grant2008dynamics}). As work gets more dependent of the retrieval and analysis of information, a proactive support system shall help the user in his various tasks by providing him with relevant information. This approach had further implications as gathering information from the user interaction with his computer requires techniques from information retrieval and computer linguistics. In this case the documents a user works with are analyzed and keywords are stored as vectors or a bag of words. The relevant keywords shall help to narrow the topical context a user is working on. Keywords than help to start searches with relevant search terms and provide the user with the information he needs (\cite{budzik2000user}). \mnote{Context-based Recommender Systems} As a single user is often not able to find the needed information, his typical search patterns are compared with those of other users. In these cases as \textit{user model} is created, and his search terms are compared to those of other users' and the documents they found. If keywords are matching, documents of those others users are recommended (\cite{anand2007contextual}). This approach is called \ac{CBRS} recommendation and their related techniques like user-collaborative filtering are applied in search engines like Amazon \footnote{www.amazon.com}.\ac{AAS} at last have different focus: The guiding principle of \acs{AAS} is that users have limited cognitive resources and are distracted easily \mnote{Attention-aware systems}. They suffer from an \textit{information overload} as they jump quickly from one resource to the next in the same and different workings tasks. Whilst it is beneficial to be able to change foci in certain situations, in others it is exhausting. Therefore systems capable of supporting and guiding user attention have to assess the current user focus, and calculate the cost/benefits of attention shifts (interruptions). As this explanation shows, \acs{AAS} have a foundation in cognitive psychology, i.e. how attention is elicited, distracted and shifting over time. Experimental setups include multiple sensor arrays like  gaze-tracking-, gesture-tracking, speech-detection and systems that measure the physiological cues (\cite{roda2006attention}). But there are also non-sensory based approaches that record users' interaction with software (\cite{horvitz2003models}, \cite{schmitz2011contextualized}). Attention management architectures expand the agenda of context-based systems, as they want not only to detect the current state of the attention of users, but also want provide support. Therefore not only the attentional state has to be tracked but the system needs to establish the users' goals and current tasks and also the happenings in the environment (\cite{roda2006attention}). Consequently this lead to the proclamation \ac{IAS}. This approach combines \acs{CAS} and \acs{AAS} by explicating individual and implicit intentions and plans of users' to reason about attention and context information. Dealing with context and attention means dealing with uncertainty. Explicated task models, so the idea, could help to increase the chances in proactive support. The term "'intention"' is  approached in the following way(\cite{cohen1990intention}):
\begin{quotation}
  Intention has often been analyzed differently from other mental states such as belief and knowledge. First, whereas the content of beliefs and knowledge is usually considered to be in the form of propositions, the content of an intention is typically regarded as an action. For example, Castefiada treats the content of an intention as a "`practition"' similar to an action description \dots. It is claimed that by doing so, and by strictly separating the logic of propositions from the logic of practitions, one avoids undesirable properties in the logic of intention, such as the fact that if one intends to do an action a one must also intend to do a or b. However, it has also been argued that needed connections between propositions and practitions may not be derivable.
\end{quotation}

The authors further argue that intention is directed towards the future actions and according plans. Intention thus shall be modeled as "'a composite concept of what an agent has chosen and how the agent is committed to that choice"'. The choice can be a desire or goal. Intention therefore can be described as a persisting goal. If intention is defined in a formal theory, then beliefs, goals and desires must be expressed in the same way. As the theory may be correct, the deductions fall short for real world problems. On the other hand, if those terms are used in a very abstract way, they can not be used for a touring machine. Either way
